%!TEX root = thesis.tex

\input{../preamble}
\addbibresource{../thesis.bib}



\title{$k$-Sided Rectangular Duals \\ \normalsize Draft 1.2}
\author{Sander Beekhuis}
\date{\today} %\today%


%
% These commands need to be defined in order to produce a correct and personalized document
%
\newcommand{\shortdoctitle}{$\mathbf{k}$-Sided Rectangular Duals}
\newcommand{\doctitle}{$\mathbf{k}$-Sided Rectangular Duals}
\newcommand{\docsubtitle}{Some catching subtitle}

\newcommand{\me}{Sander Beekhuis}
\newcommand{\keywords}{keyword1, keyword2, keyword3}
\newcommand{\version}{Version Draft 1.3.7}
\newcommand{\monthYear}{Feb 2017}

%Be sure to use all the titles for your committee members!!! (their names show up on the very first page!)
\newcommand{\firstCommitteeMember}{Your First Committee Member}
\newcommand{\secondCommitteeMember}{Your Second Committee Member, usually the daily supervisor}
\newcommand{\thirdCommitteeMember}{Your Third Committee Member, usually the external member}

\author{\me}

%
% PDF settings
%
\hypersetup
{
    pdfauthor={\me},
    pdftitle={\shortdoctitle},
    pdfsubject={\doctitle},
    pdfkeywords={\keywords}
}


\begin{document}
\pagenumbering{roman}



\include{titlepage}
\normalsize

%todos
\newpage
\fxinnote{Align figures and tables}

\fxinnote{Q: What is the right order of my committee members? I can't recall what order we decided.}

\fxinnote{Q: Think about title and subtitle}

\fxinnote{Do something with keywords or drop them}

\fxinnote{Q: I should/could do running time analysis.}

\fxinnote{Q: How do I organize all subsubsection for the sweepcycle subsection? }

\fxinnote{Q: What is a preface? What should be in there? Should I have one? }


%ABSTRACT
\newpage
\include{abstract/abstract}

%PREFACE
\newpage
\thispagestyle{plain}

\tableofcontents

\newpage
\thispagestyle{plain}

\listoffigures

\newpage
\thispagestyle{plain}

\listoftables
\fxnote{Q: This looks silly. Drop list of tables?}

\newpage
\setcounter{page}{0}
\pagenumbering{arabic}
%from here on, start the 'real' page numbering, from 1, with normal digits


\include{introduction/introduction}
%\include{prelim/prelim}

\include{rectangularDuals/rectangularDuals}

\include{fixExtension/fixExtension}

\include{unifiedAlgo/algo}

\include{unifiedAlgo/rightNeighbourWalk}

\include{unifiedAlgo/sweepCycle}

\include{unifiedAlgo/flipBlueZ}

\include{topFanFlips/topFanFlips}

\include{blueFaceSubdivision/blueFaceSubdivision}

%\include{examples/examples}

%\section{Running time analysis}
\thispagestyle{plain}
%\section{Conclusion}
\thispagestyle{plain}


\listoffixmes

%\include{unusedTheorems/connectedTriangulations}
%\include{unusedTheorems/unusedTheorems}
%\include{conclusion/researchQuestions}

\newpage
\thispagestyle{plain}


\printbibliography
\end{document}
